---
title: "Exercise Manner Prediction"
author: "Matthew May"
date: "2018-2-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## abstract
This report provides an prediction model based on the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. It's goal is to predict the manner in which they did the exercise. I use random forest as the model for this task. I also use cross validation during trainning. The accuracy is 99.56%.

## clean the data
Since there are many NA in the data, I have to clean the data first  

```{r}


set.seed(1)

data <- read.csv("D:\\coursera\\8Machine learning\\week4\\Peer-graded Assignment Prediction Assignment Writeup\\pml-training.csv")
data <- data[,c("roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x","gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z","magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm","gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell","pitch_dumbbell","yaw_dumbbell","gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z","roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm","gyros_forearm_x","gyros_forearm_y","gyros_forearm_z","accel_forearm_x","accel_forearm_y","accel_forearm_z","magnet_forearm_x","magnet_forearm_y","magnet_forearm_z","classe")]

```


## build model

* I split the data into training set and testing set. 
* I create the trainControl object to do the cross-validation with 5 folds
* I have tried several different models and random forest is proved to be the best, so I use random forest here. 
  
``` {r}
library("caret")
inTrain <- createDataPartition(data$classe,p=0.7,list=FALSE)

training <- data[inTrain,]
testing <- data[-inTrain,]


# training----------------------------------------------------------------------
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)

fitControl <- trainControl(method = "cv",
                           number = 5,
                           allowParallel = TRUE)


fit <- train(classe ~ .,method="rf",training,trControl=fitControl)

stopCluster(cluster)
registerDoSEQ()
```




## testing and out of sample error
For discrete data, we use 'accuracy' to measure the performance.  
The test shows that the accuracy is 0.995582  
  
```{r}
p <- predict(fit,testing)

# Accuracy : 0.9956
confusionMatrix(p,testing$classe)

```

## predict on the test.csv
``` {r}
d_test <- read.csv("D:\\coursera\\8Machine learning\\week4\\Peer-graded Assignment Prediction Assignment Writeup\\pml-testing.csv")
d_test <- d_test[,c("roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x","gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z","magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm","gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell","pitch_dumbbell","yaw_dumbbell","gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z","roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm","gyros_forearm_x","gyros_forearm_y","gyros_forearm_z","accel_forearm_x","accel_forearm_y","accel_forearm_z","magnet_forearm_x","magnet_forearm_y","magnet_forearm_z")]
p <- predict(fit,d_test)
```


